Windowed least-squares model reduction for dynamical systems

### 2019-10-18 main.tex
update language in proof1
clean up and conclusions

\section{Conclusions}\label{sec:conclude}
This paper introduced the windowed least-squares (\methodAcronym) \approachKwd\ for model reduction  of dynamical systems. The approach sequentially minimizes the time-continuous full-order model residual within a low-dimensional trial subspace over a series of time windows. The formulation is applicable to both spatial and space--time dimension reduction trial subspaces. The Galerkin, Least-Squares Petrov--Galerkin (LSPG) and space--time LSPG methods were shown to be specific instances of the proposed method

We considered two differing techniques to solve the least-squares minimization problems arising from the \methodAcronym\ \approachKwd: direct (discretize-then-optimize) and indirect (optimize-then-discretize). 
The case of \spatialAcronym\ trial subspaces proved particularly interesting: it was shown that indirect methods comprise solving a coupled two-point 
Hamiltonian boundary value problem. The forward problem evolves the standard Galerkin ROM forced by a ``costate" variable, while a backward problem evolves the costate and is forced by the ROM residual.

Numerical experiments of the compressible Euler and Navier--Stokes equations demonstrated the utility in the proposed approach. The first numerical experiment, in where the Sod shock tube was examined, demonstrated that \methodAcronymROMs\ minimizing the residual over larger time windows yielded solutions with lower space--time residuals. Increasing the window size over which the residual was minimized, however, did not necessarily decrease the $\elltwo$ solution error; this was observed to occur over an intermediary window size. We additionally observed that the  
\methodAcronym\ \approachKwd\ overcomes the time-step sensitivity challenges that LSPG is subject to. The second numerical experiment, which examined collocated ROMs of a compressible cavity flow, demonstrated the utility of the \methodAcronym\ formulation on a more complex flow. In this experiment, \methodAcronymROMs\ yielded predictions with relative errors less than $5-10\%$, while the Galerkin and LSPG ROMs failed to converge/blew up. Increasing the window size over which the residual was minimized again led to a lower space--time residual, but not necessarily a lower $\elltwo$ error.

@@

The principal challenge encountered in the \methodAcronym\ formulation is the computational cost: increasing the window size over which the residual is minimized leads to, in general, a higher computational cost. In the context of the direct solution approach, this increased cost is due to the fact that forming and solving the least-squares problem becomes more expensive as the window size grows. In the context of indirect methods, this increased cost is a result of the increased number of iterations of the forward backward sweep method. While numerical experiments demonstrated that the increase in computational cost is mild, future work will focus on the development of solution techniques to improve the scaling of the method.

@@

	$\statez(a) = \statez_a$. \KTC{This previous sentence isn't correct. The
	functional defined here has two arguments; the extremum can't be a single
	argument, as it's a pair of arguments. You need to connect these by defining
	a new function that maps a time-dependent function to this functional
	evaluated at the time-dependent function and its derivative..}\EP{ Check this, I think this is more precise now.} 
We introduce an arbitrary function $\variation : [a,b] \rightarrow \RR{M}$

@@

We define a new function that is equivalent to~\eqref{eq:el_gen} evaluated at $\statezBar$ in the first argument and 

@@

\begin{align}\label{eq:gradient}

@@

	=&  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete }

@@

	,t \big) \bigg) \variation(t) \bigg]dt + \\ &\frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b,\epsilon) ,b \big) \variation(b) , 

@@

Substituting Eq.~\eqref{eq:gradient} in the stationarity conditions~\eqref{eq:stationaryOne} yields

@@

\begin{equation}

@@

\begin{gather}\label{eq:euler_lagrange_appendix}
\bigg[ \frac{\partial \integrand  }{\partial \stateyDiscrete }

@@

	\big(\statez(t),\statezDot(t),t \big) \bigg) \bigg]^T = \bz  \\
\statez(a) = \statez_a, \qquad \bigg[ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(b),\statezDot(b),b \big) \bigg]^T = \bz.
\end{gather}

@@

\section{Evaluation of gradients in the Euler--Lagrange equations for \methodAcronym\ with \spatialAcronym\ trial subspaces}\label{appendix:vector_calc}
We now derive the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation with \spatialAcronym\ trial subspaces. Without loss of generality, we present the derivation for a single window $t \in [0,T]$ with a constant basis $\basisspace$ and weighting matrix $\stweightingMat$.  
To obtain the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation, we need to evaluate the gradients in~\eqref{eq:euler_lagrange_appendix} for the integrand
\begin{align*}\label{eq:integrand_apx}
 \minintegrand & \vcentcolon
(\genstateyDiscreteArgnt{}, \genstateyDiscreteDotArgnt{} ,\timeDummy) \mapsto \frac{1}{2} \big[
\basisspace \genstateyDiscreteDotArgnt{} - \velocity(\basisspace \genstateyDiscreteArgnt{}
+ \stateIntercept,\timeDummy) \big]^T \stweightingMatArg{} \big[
\basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} +
\stateIntercept,t) \big], \\ & 
\vcentcolon \RR{\romdim} \times \RR{\romdim} \times [0,T]
 \rightarrow \RR{} .  
\end{align*}
To evaluate the gradients, we first expand $\minintegrand$:
\begin{align*}
 \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy)  &= \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} + \stateIntercept,\timeDummy) \big]^T \stweightingMatArg{} \big[ \basisspace \genstateyDiscreteDotArgnt{} - \velocity(\veloargsromy) \big] \\ 
 &= \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[\basisspace \genstateyDiscreteDotArgnt{} \big]  - \frac{1}{2}\big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{}  \big[\basisspace \genstateyDiscreteDotArgnt{}\big] - \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  \\ &+ \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big].
\end{align*}
Since $\stweightingMat$ is symmetric,
\begin{equation}\label{eq:int_expand}
 \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy)  = \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  + \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big].
\end{equation}
For notational purposes, we write the above as,
$$  \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) =  \minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) +  \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) +  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy),$$
where
\begin{align*}
&\minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv  \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big], \\
&   \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv   -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big], \\
&  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) \equiv  \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] . 
\end{align*}
Constructing the Euler--Lagrange equations for this functional $\minintegrand$ requires evaluating the derivatives $\frac{\partial \minintegrand}{\partial \genstateyDiscrete}$ and $\frac{\partial \minintegrand}{\partial \genstateyDiscreteDot}$. We start by evaluating $\frac{\partial \minintegrand}{\partial \genstateyDiscrete}$ and go term by term.

>>>>

\section{Conclusions}\label{sec:conclude}
This paper proposed the windowed least-squares (\methodAcronym) \approachKwd\ for model reduction  of dynamical systems. The approach sequentially minimizes the
	time-continuous full-order-model residual within a low-dimensional space--time trial
	subspace over time windows. The approach was formulated for two types of trial subspaces: one that reduces only the spatial dimension of the full-order model, and one that reduces both the spatial and temporal dimensions of the full-order model. For each type of trial
	subspace, we then outlined two different solution techniques: direct (i.e.,
	discretize then optimize) and indirect (i.e., optimize then discretize). We showed that particular instances of the approach recover Galerkin,
	least-squares Petrov--Galerkin (LSPG), and space–time LSPG projection.

The case of \spatialAcronym\ trial subspaces is of particular interest in the \methodAcronym\ context: it was shown that indirect methods comprise solving a coupled two-point 
Hamiltonian boundary value problem.
The forward system, which is forced by an auxiliary costate,
		evolves the (spatial) generalized coordinates of the ROM in time. The
		backward system, which is forced by the time-continuous FOM residual
		evaluated about the ROM state, governs the dynamics of the costate.

Numerical experiments of the compressible Euler and Navier--Stokes equations demonstrated the utility in the proposed approach. The first numerical experiment, in where the Sod shock tube was examined, demonstrated that \methodAcronymROMs\ minimizing the residual over larger time windows yielded solutions with lower space--time residuals. Increasing the window size over which the residual was minimized, however, did not necessarily decrease the solution error in the $\elltwo$-norm; this was observed to occur over an intermediary window size. We additionally observed that the  
\methodAcronym\ \approachKwd\ overcomes the time-discretization sensitivity that LSPG is subject to. The second numerical experiment, which examined collocated ROMs of a compressible cavity flow, demonstrated the utility of the \methodAcronym\ formulation on a more complex flow. In this experiment, \methodAcronymROMs\ yielded predictions with relative errors less than $5-10\%$, while the Galerkin and LSPG ROMs failed to converge/blew up. Increasing the window size over which the residual was minimized again led to a lower space--time residual, but not necessarily a lower error in the $\elltwo$-norm.

@@

The principal challenge encountered in the \methodAcronym\ formulation is the computational cost: increasing the window size over which the residual is minimized leads to a higher computational cost. In the context of the direct solution approach, this increased cost is due to the increased cost of forming and solving the least-squares problem associated with larger window sizes. In the context of indirect methods, this increased cost is a result of the increased number of iterations of the forward backward sweep method. While numerical experiments demonstrated that the increase in computational cost is mild, future work will focus on the development of solution techniques to improve the scaling of the method.

@@

	$\statez(a) = \statez_a$. 
We additionally introduce an arbitrary function $\variation : [a,b] \rightarrow \RR{M}$

@@

We define a new function that is equivalent to the function~\eqref{eq:el_gen} evaluated at $\statezBar$ in the first argument and 

@@

\begin{multline}\label{eq:gradient}
%	\begin{split}

@@

	=  \int_a^b  \bigg[\frac{\partial \integrand  }{\partial \stateyDiscrete }

@@

	,t \big) \bigg) \variation(t) \bigg]dt + \\ 
\frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statezBar(b,\epsilon),\statezDotBar(b,\epsilon) ,b \big) \variation(b) , 
%	\end{split}
\end{multline}

@@

Substituting Eq.~\eqref{eq:gradient} in the stationarity condition~\eqref{eq:stationaryOne} yields

@@

\begin{equation}\label{eq:euler_lagrange_analysis}

@@

\begin{equation}\label{eq:euler_lagrange_appendix}
\bigg[ &\frac{\partial \integrand  }{\partial \stateyDiscrete }

@@

	\big(\statez(t),\statezDot(t),t \big) \bigg) \bigg]^T = \bz , \\
&\statez(a) = \statez_a, \hspace{0.65 in} \bigg[ \frac{\partial \integrand}{\partial \stateyDotDiscrete} \big(\statez(b),\statezDot(b),b \big) \bigg]^T = \bz.

@@

\section{Evaluation of gradients in the Euler--Lagrange equations for \methodAcronym\ with \spatialAcronym\ trial subspaces}\label{appendix:vector_calc}
We now derive the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation with \spatialAcronym\ trial subspaces. Without loss of generality, we present the derivation for a single window $t \in [0,T]$ with a constant basis $\basisspace$ and weighting matrix $\stweightingMat$.  
To obtain the specific form of the Euler--Lagrange equations for the \methodAcronym\ formulation, we need to evaluate the gradients in~\eqref{eq:euler_lagrange_appendix} for the integrand
\begin{align*}\label{eq:integrand_apx}
 \minintegrand & \vcentcolon
(\genstateyDiscreteArgnt{}, \genstateyDiscreteDotArgnt{} ,\timeDummy) \mapsto \frac{1}{2} \big[
\basisspace \genstateyDiscreteDotArgnt{} - \velocity(\basisspace \genstateyDiscreteArgnt{}
+ \stateIntercept,\timeDummy) \big]^T \stweightingMatArg{} \big[
\basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} +
\stateIntercept,\timeDummy) \big], \\ & 
\vcentcolon \RR{\romdim} \times \RR{\romdim} \times [0,T]
 \rightarrow \RR{} .  
\end{align*}
To evaluate the gradients, we first expand $\minintegrand$:
\begin{align*}
 \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy)  &= \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{}  - \velocity(\basisspace \genstateyDiscreteArgnt{} + \stateIntercept,\timeDummy) \big]^T \stweightingMatArg{} \big[ \basisspace \genstateyDiscreteDotArgnt{} - \velocity(\veloargsromy) \big] \\ 
 &= \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[\basisspace \genstateyDiscreteDotArgnt{} \big]  - \frac{1}{2}\big[ \velocity(\veloargsromy) \big]^T \stweightingMatArg{}  \big[\basisspace \genstateyDiscreteDotArgnt{}\big] - \frac{1}{2} \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  \\ &+ \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big].
\end{align*}
Since $\stweightingMat$ is symmetric,
\begin{equation}\label{eq:int_expand}
 \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy)  = \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]  -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big]  + \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big].
\end{equation}
For notational purposes, we write the above as,
$$  \minintegrand(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) =  \minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) +  \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) +  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy),$$
where
\begin{align*}
&\minintegrand_1(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) =  \frac{1}{2}\big[\basisspace \genstateyDiscreteDotArgnt{}  \big]^T  \stweightingMatArg{}  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big], \\
&   \minintegrand_2(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) = -  \big[ \basisspace \genstateyDiscreteDotArgnt{} \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big], \\
&  \minintegrand_3(\genstateyDiscrete,\genstateyDiscreteDot,\timeDummy) = \frac{1}{2}\big[\velocity(\veloargsromy) \big]^T \stweightingMatArg{} \big[ \velocity(\veloargsromy) \big] . 
\end{align*}
Constructing the Euler--Lagrange equations for this functional $\minintegrand$ requires evaluating the derivatives $\frac{\partial \minintegrand}{\partial \genstateyDiscrete}$ and $\frac{\partial \minintegrand}{\partial \genstateyDiscreteDot}$. We start by evaluating $\frac{\partial \minintegrand}{\partial \genstateyDiscrete}$ and go term by term.

### 2019-10-18 section4/spacetime_numerical.tex
clean up and conclusions

\item replace $J_{D-ST}^n$ in (38) with $J_D^n$. You can do this because of Eq 
	(33) \EP{Yes, done}
\item remove definition of $J_{D-ST}$. \EP{Yes, done}

@@

\EP{I noticed you got rid of the assumption about the size of the generalized coordinates. Why don't we need to worry about this?}

@@

The boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace. The minimization problem~\eqref{eq:obj_gen_lms_final_st} again yields a least-squares problem.

@@

For the limiting case where one window comprises entire domain (i.e.,

@@

	space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrix $\lspgWeightingST = \text{diag}(\stweightingMatOneArg{1})$

@@

	would be more general.}\EP{Yes, this is better. Is diag the right command though? It should be $W_{ST} = [W \; W \; W \; W]$}, and $\nstepsArg{1} = N_t$ time instances are employed that

@@

\bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg] \\

@@

\KTC{It would be great to a comment about how this is equivalent to Qiqi Wang
and Paul Constantine's method in the case of one time window and $\Pi^1$
corresponding to full solution trajectories. They use collocation-based
hyper-reduction at a few points in time (but for the full ODE residual at that
time), which you could mention. This is a very cool connection I didn't
realize before, and I think it's definitely worth mentioning!}
%$$\objectiveDiscreteSTArg{n}: \stgenstatey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 
%\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptSTArg{n},\collocPointSTArg{n,i}) \bigg]^T 
%\stweightingMatArg{n} 
%\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptSTArg{n},\collocPointSTArg{n,i}) \bigg].
%The stationary points are given by,
%\begin{multline*}\label{eq:st_stationary_discrete}
%\frac{\partial \objectiveDiscreteSTArg{n}}{\partial \stgenstatey}(\stgenstateArg{n}) = 
%\sum_{i=0}^{\ncollocSTArg{n}} \gamma_i \bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}}^T  - \stbasisArg{n}{\collocPointSTArg{n}{i}}^T \bigg[\frac{\partial
%\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n} +                    
%\stateInterceptArg{n})\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n} + \stateInterceptArg{n}) \bigg),\end{multline*}
%%subject to the boundary conditions 
%\begin{equation}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\approxstateIC &
%n=1. \end{cases} \end{equation} 
\subsection{\spaceTimeAcronym\ trial subspaces: summary}
\spaceTimeAcronym\ trial subspaces yield a series of space--time systems of algebraic equations over each window. As a variety of work has examined space--time reduced-order models with \spaceTimeAcronym\ trial subspaces, 
a detailed exposition of solution techniques for these systems is not pursued here. It is sufficient to say that the 
space--time trial subspace yields a series of dense systems to be solved over each window.

>>>>

@@

\EP{I noticed you got rid of the assumption about the size of the generalized coordinates. I think we need to mention this, I put in a note below. Let me know if I'm missing something}

@@

The boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace. 
Assuming $\text{Rank}(\stweightingMatOneArg{n}) \nstepsArg{n} \ge \stdimArg{n}$, the minimization problem~\eqref{eq:obj_gen_lms_final_st} again yields a least-squares problem.

@@

For the limiting case where one window comprises the entire domain (i.e.,

@@

	space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrix $\lspgWeightingST = \begin{bmatrix} \stweightingMatOneArg{1} & \cdots & \stweightingMatOneArg{1}\end{bmatrix}$

@@

	would be more general.}\EP{Yes, this is better. Not sure if diag is right though, see if you agree.}, and $\nstepsArg{1} = N_t$ time instances are employed that

@@

\bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg], \\

@@

\begin{remark}
\methodAcronym\ with \spaceTimeAcronym\ trial subspaces solved via the indirect approach naturally achieves ``collocation" in time as the full-order model residual needs to be queried at only the quadrature points. 
\end{remark}
\begin{remark}
For the limiting case where one window comprises the entire domain (i.e.,
	$\DeltaSlabArg{1}\equiv T$), the trial
	space is set to comprise the span of full solution trajectories, $\stspaceSTArg{1} = \text{span}\{ \stateFOMArgnt{}_i\}_{i=1}^{\stdimArg{1}}$ (where $\stateFOMArgnt{}_i$ are obtained, e.g.,  from training simulations at different parameter instances), and the weighting matrix is set to $\stweightingMatArg{1} = \mathbf{I}$, \methodAcronym\ with \spaceTimeAcronym\ trial subspaces and the indirect approach closely resembles the model reduction procedure proposed in~\cite{constantine_strom}; the approaches differ only in that Ref.~\cite{constantine_strom} places the constraint $\sum_{i=1}^{\stdimArg{1}} \stgenstateArg{1}_i = 1$. 
\end{remark}

\KTC{It would be great to a comment about how this is equivalent to Qiqi Wang
and Paul Constantine's method in the case of one time window and $\Pi^1$
corresponding to full solution trajectories. They use collocation-based
hyper-reduction at a few points in time (but for the full ODE residual at that
time), which you could mention. This is a very cool connection I didn't
realize before, and I think it's definitely worth mentioning!}
\EP{See if you agree with what I have. I worked with the trial space instead of the basis because we prescribe certain structure to the basis (such as the affine transformation)}.
%$$\objectiveDiscreteSTArg{n}: \stgenstatey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 
%\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptSTArg{n},\collocPointSTArg{n,i}) \bigg]^T 
%\stweightingMatArg{n} 
%\bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ }  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateyArg{ } + \stateInterceptSTArg{n},\collocPointSTArg{n,i}) \bigg].
%$$
%The stationary points are given by,
%\begin{multline*}\label{eq:st_stationary_discrete}
%\frac{\partial \objectiveDiscreteSTArg{n}}{\partial \stgenstatey}(\stgenstateArg{n}) = 
%\sum_{i=0}^{\ncollocSTArg{n}} \gamma_i \bigg[ \stbasisDotArg{n}{\collocPointSTArg{n}{i}}^T  - \stbasisArg{n}{\collocPointSTArg{n}{i}}^T \bigg[\frac{\partial
%\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n} +                    
%\stateInterceptArg{n})\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{\collocPointSTArg{n}{i}} \stgenstateArg{n} + \stateInterceptArg{n}) \bigg),\end{multline*}
%%subject to the boundary conditions 
%\begin{equation}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\approxstateIC &
%n=1. \end{cases} \end{equation} 
\subsection{\spaceTimeAcronym\ trial subspaces: summary}
\spaceTimeAcronym\ trial subspaces yield a series of space--time systems of algebraic equations over each window. As a variety of work has examined space--time reduced-order models with \spaceTimeAcronym\ trial subspaces, 
a detailed exposition of solution techniques for these systems is not pursued here. It is sufficient to say that \methodAcronym\ with \spaceTimeAcronym\ trial subspaces yields a series of dense systems to be solved over each window.
